# Gapminder Dataset
This dataset has 6 variables and 3,313 records. This dataset is from R built-in
package gapminder.

# Questions
1. Which Continent has the highest pop, gdpPercap & lifeExp?
2. Does pop, lifeExp, gdpPercap increases with passing years?

Packages Required

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(gapminder)
library(ggplot2)
library(gridExtra)
library(plotly)
```

Loading dataset into variable df.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# load the data fom gapminder package
df <- gapminder_unfiltered

# looking at the dataset
str(df)
```

Variables are
1. country - representing the country's data
2. continent - represents the continents countries belong to.
3. year - is the year for which the data was takes for that particular country.
4. lifeExp - life expectancy in years, of a person born at that given time.
5. pop - population is the population of the country.
6. gdpPercap - is gross domestic product divided by population.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# checking for NA's
sum(is.na(df))
```

There're no missing records in the dataset.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# viewing first few records of the dataset
head(df)
```

Few records from the dataset showing the data.

# Univariate Analysis

```{r echo=FALSE, message=FALSE, warning=FALSE}
# plotting histogram for lifeExp
h1 <- ggplot(data = df, aes(x = lifeExp)) +
  geom_histogram(binwidth = 1) +
  ggtitle('Life Expectancy Histogram')

# plotting histogram for gdpPercap
h2 <- ggplot(data = df, aes(x = gdpPercap)) +
  geom_histogram(binwidth = 1000) +
  ggtitle('gdpPercap Histogram')

grid.arrange(h1, h2, ncol = 2)
```

These histograms show the distribution of gdpPercap & lifeExp variables.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# bar chart for coninents
br1 <- ggplot(data = df, aes(x = continent)) +
  geom_bar() +
  ggtitle('Continent Bar Chart')

grid.arrange(br1, ncol = 1)
```

Europe has highest number of records in the dataset.

# Bivariate Analysis

## Question 1 - Which Continent has the highest pop, gdpPercap & lifeExp?

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Boxplot for continents with life expectancy
b1 <- ggplot(data = df, aes(x = continent, y = lifeExp)) +
  geom_jitter(alpha = 0.3) +
  geom_boxplot(alpha = 0.5, outlier.colour = "blue") +
  stat_summary(fun.y = "mean", color = "red", shape = 8, geom = "point") +
  ggtitle('Continent wise Life Expectancy')

grid.arrange(b1, ncol = 1)
```

This boxplot depicts that Europe has the highest life expectancy.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Boxplot for continents with GDP per capita
b2 <- ggplot(data = df, aes(x = continent, y = gdpPercap)) +
  geom_jitter(alpha = 0.3) +
  geom_boxplot(alpha = 0.5, outlier.colour = "blue") +
  stat_summary(fun.y = "mean", color = "red", shape = 8, geom = "point") +
  coord_cartesian(ylim = c(240, 40000)) +
  ggtitle('Continent wise gdpPercap')

grid.arrange(b2, ncol = 1)
```

This boxplot depicts that Europe has the highest GDP Per Capita.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Boxplot for continents with population
b3 <- ggplot(data = df, aes(x = continent, y = pop)) +
  geom_jitter(alpha = 0.3) +
  geom_boxplot(alpha = 0.5, outlier.colour = "blue") +
  stat_summary(fun.y = "mean", color = "red", shape = 8, geom = "point") +
  ggtitle('Continent wise population')

grid.arrange(b3, ncol = 1)
```

Asia has the highest population.

```{r echo=FALSE, message=FALSE, warning=FALSE}
cm <- ggcorr(df[, 3:6], label = TRUE, nbreaks = 4, method = c("all.obs", "spearman"),
       hjust=0.8, angle=-70, size=4) +
  ggtitle('Spearman Correlation Coefficient Matrix')

grid.arrange(cm, ncol = 1)
```

Looking at the coefficient correlation matrix, lifeExp & gdpPercap has the 
strongest correlation.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# scatter plot for lifeExp vs gdpPercap
sp1 <- ggplot(data = df, aes(x = lifeExp, y = gdpPercap)) +
  geom_jitter(alpha = 0.1, color = "orange") +
  geom_smooth(se = TRUE) +
  ggtitle('lifeExp vs gdpPercap Scatter Plot')

grid.arrange(sp1, ncol = 1)
```

This is the strongest relation and gdp increases if life expectancy is 60 
or greater.

## Question 2 - Does lifeExp, gdpPercap increases with passing years?

```{r echo=FALSE, message=FALSE, warning=FALSE}
# scatter plot for year vs lifeExp
sp2 <- ggplot(data = df, aes(x = year, y = lifeExp)) +
  geom_jitter(alpha = 0.1, color = "orange") +
  geom_smooth(se = TRUE) +
  ggtitle('year vs lifeExp Scatter Plot')

grid.arrange(sp2, ncol = 1)
```

This plot depicts with passing years Life Expectancy is increasing.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# scatter plot for lifeExp vs gdpPercap
sp3 <- ggplot(data = df, aes(x = year, y = gdpPercap)) +
  geom_jitter(alpha = 0.1, color = "orange") +
  geom_smooth(se = TRUE) +
  scale_y_continuous(limits = c(0,
                                quantile(df$gdpPercap, probs = 0.95))) +
  ggtitle('year vs gdpPercap Scatter Plot')

grid.arrange(sp3, ncol = 1)
```

Here looking at the plot, looks that gdpPercap has weak correlation with years.

# Multivariate Analysis

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Plotting bubble chart only for year 2007
m <- df %>%
  filter(year == 2007) %>%
  ggplot(aes(x = gdpPercap, y = lifeExp, size = pop, colour = continent)) +
  geom_point() +
  scale_x_log10() +
  theme_bw() +
  ggtitle("Bubble Chart For Year 2007 gdpPerCap vs lifeExp")

f <- ggplotly(m)

```

This bubble chart is plotted for year 2007, for all the continents.
Looking at this bubble chart, Europe & Americas have maximum gdpPercap with 
maximum lifeExp. And it seems like the size of population does not affect the
gdpPercap.
